Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa through development of a 3D spectral numerical coastal-flow model

### 2020-08-05 thesis_chapter_1.Rmd
added heading
added to drag section

Drag is the force an object experiences when moving in a fluid due to the rate in change of momentum [@Vogel2020]. The drag around the surface of an object due to the viscosity of the fluid is known as 'surface drag'. The drag induced by a quantity of fluid moving around the object is known as pressure drag/form drag/inertial drag. In general, large objects moving quickly tend to induce larger drag forces, however, this is not always the case. In some instances, the flow around the object changes from laminar to turbulent which ultimately reduces overall drag. (Maybe insert small section on why turbulent flow reduces overall drag on an object).

The type of flow, laminar or turbulent, is determined by fluid density (p), flow speed (U), object size (l) and fluid viscosity (v). These factors can be used to calculate the Reynolds number which is used to determine the type of flow. The expression used can be dependent on factors used such as whether the fluid is Newtonian or non-Newtonian and object shape.

@@

The trajectory of kelp rafts are determined by a combination of oceanographic conditions and biological processes. Since kelp  float, evidence points largely to wind driven surface currents as the main driver, although the relative importance of wind versus surface currents is largely unknown. For example a study by @harrold1989 used radio-trackers on both natural and artificial *Macrocystis pyrifera* kelp rafts to investigate seasonal trajectory patterns. The results showed that tracked kelp rafts were mostly deposited nearby the source population and that seasonal wind direction was the primary driver of trajectory for kelp rafts submerged roughly ~0.5 - 1m below the surface @harrold1989. However, the authors did note the influence of surface currents for some of their experiments when winds relaxed; which suggests that surface currents may play a larger role in determining trajectory for kelp rafts less exposed to wind or when prevailing winds are low. Other authors have identified the West Wind Drift as important driver of trajectory of floating *M. pyrifera* [@Fraser2011; ]

>>>>

Drag is the force an object experiences when moving in a fluid due to the rate in change of momentum [@Vogel2020]. The drag around the surface of an object due to the viscosity of the fluid is known as 'surface drag'. The drag induced by a quantity of fluid moving around the object is known as pressure drag/form drag/inertial drag. In general, large objects moving quickly tend to induce larger drag forces, however, this is not always the case. In some instances, the flow around the object changes from laminar to turbulent which ultimately reduces overall drag. (Maybe insert small section on why turbulent flow reduces overall drag on an object). The type of flow, laminar or turbulent, is determined by fluid density (p), flow speed (U), object size (l) and fluid viscosity (v). These factors can be used to calculate the Reynolds number which is used to determine the type of flow. The expression used can be dependent on factors used such as whether the fluid is Newtonian or non-Newtonian and object shape.

The drag coefficient is the Fluid Dynamic Drag or is also known as the Hydrodynamic Drag, but from here on it will be referred to simply as Drag. The drag is determined by two factors, which may also be considered to be sub-components of drag. These factors, or sub-components, are Skin Friction and Form Drag and are partly determined by characteristics of the object and the medium in which the object is moving. For instance, skin drag is determined by the “roughness” of the object’s surface and the form drag by the shape and size of the object. An object which is moving through a liquid vertically will have multiple forces acting on it, namely the viscosity coefficient of the liquid $(F_A)$, the buoyancy force $(BF)$ and the Drag $(F_D)$; provided the object’s density is greater than that of the density of the liquid. The viscosity coefficient is dependent on the characteristics of the medium and the buoyancy force by the mass of the object and the gravitational forces acting on it. For the object in this example to move downwards, some of the liquid needs to be displaced away from the object which requires a force. This force gives the liquid a moment of inertia and is dependent on the shape of the object and ultimately the drag force acting on the object as it moves through the liquid. As an example, a short cylinder has less drag than that of a long cylinder as there is more liquid that needs to be displaced for the object to move through the medium.

@@

The trajectory of kelp rafts are determined by a combination of oceanographic conditions and biological processes. Since kelp  float, evidence points largely to wind driven surface currents as the main driver, although the relative importance of wind versus surface currents is largely unknown. For example a study by @harrold1989 used radio-trackers on both natural and artificial *Macrocystis pyrifera* kelp rafts to investigate seasonal trajectory patterns. The results showed that tracked kelp rafts were mostly deposited nearby the source population and that seasonal wind direction was the primary driver of trajectory for kelp rafts submerged roughly ~0.5 - 1m below the surface @harrold1989. However, the authors did note the influence of surface currents for some of their experiments when winds relaxed; which suggests that surface currents may play a larger role in determining trajectory for kelp rafts less exposed to wind or when prevailing winds are low. Other authors have identified the West Wind Drift as important driver of trajectory of floating *M. pyrifera* [@Fraser2011].
